The K2 Bright Star Survey I: Methodology and Data Release

### 2019-07-11 paper/ms.tex
yu affiliation
aerts edits

\author[0000-0002-0007-6211]{Jie Yu}
\affiliation{Sydney Institute for Astronomy, School of Physics A28, The University of Sydney, NSW 2006, Australia}
\affiliation{Stellar Astrophysics Centre, Department of Physics and Astronomy, Aarhus University, DK-8000 Aarhus C, Denmark}

@@

\author[0000-0003-1822-7126]{Conny Aerts}
\affiliation{Instituut voor Sterrenkunde, KU Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium}
\affiliation{Department of Astrophysics, IMAPP, Radboud University Nijmegen, P.O. Box 9010, NL-6500 GL Nijmegen, The Netherlands}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
While the \kepler\ Mission was designed to look at tens of thousands of faint stars ($V \gtrsim 12$), brighter stars that saturated the detector are important because they can be and have been observed very accurately by other instruments. By analyzing the unsaturated scattered-light `halo' around these stars, we have retrieved precise light curves of most of the brightest stars in \ktwo fields from Campaign~4 onwards. The halo method does not depend on the detailed cause and form of systematics, and we show that it is effective at extracting light curves from both normal and saturated stars. The key methodology is to optimize the weights of a linear combination of pixel time series with respect to an objective function. We test a range of such objective functions, finding that \textit{lagged Total Variation}, a generalization of Total Variation, performs well on both saturated and unsaturated \ktwo targets. Applying this to the bright stars across the \ktwo Campaigns reveals stellar variability ubiquitously, including effects of stellar pulsation, rotation, and binarity. We describe our pipeline and present a catalogue of the 161 bright stars, with classifications of their variability, asterosesmic parameters for giants with well-measured solar-like oscillations, and remarks on interesting objects. These light curves are publicly available as a High Level Science Product from the Mikulski Archive for Space Telescopes (MAST). \href{https://github.com/benjaminpope/k2halo}{\color{linkcolor}\faGithub} % git link shamelessly borrowed from Luger!
\end{abstract}

@@

% intro kepler
The \kepler Space Telescope was launched with a main goal of determining the frequency of Earth-sized planets around Solar-like stars \citep{2010sci...327..977b}. In order to explore these populations it was necessary to observe hundreds of thousands of stars, with the consequence that the \kepler exposure time and gain were set to optimally observe eleventh or twelfth-magnitude stars, while bright stars are saturated and intentionally avoided. In the two-wheeled revival as the \ktwo mission, the \kepler telescope observed a sequence of ecliptic-plane fields containing many more very-saturated stars \citep{2014PASP..126..398H}. While it is difficult to obtain precise light curves of these stars because of their saturation, they are some of the most valuable targets to follow up with photon-hungry methods such as interferometry and high-resolution spectroscopy, and they typically have long histories of previous observations. Dedicated bright-star space photometry missions such as MOST \citep{most} and the BRITE-Constellation \citep{brite,brite2} use very small telescopes (15 and~20\,cm apertures respectively), and we would prefer to use much larger telescopes such as \kepler (0.95\,m) to obtain higher precision light curves.

@@

% describe halo algorithm
The `TV-min' halo method was first described by \citet{White2017} and applied to the Pleiades' Seven Sisters. It was also applied to Aldebaran with further developments by \citet{Farr2018}. We follow the `OWL' concept described by Hogg \&  Foreman-Mackey (2014, unpublished: preprint  \href{https://github.com/davidwhogg/OWL/}{\nolinkurl{github.com/davidwhogg/OWL/}}) in our assumptions. We assume that a star has a wide PSF sampled by many pixels with different sensitivities, which moves around on the detector within a small region and which may vary to a small extent in time. We assume that our time series consists of many epochs sampled with a nearly even cadence. We do not have metadata describing the spacecraft motion, pixel gains, PSF variations or other noise processes. Because photometry is a linear operation, any estimator of the flux is necessarily a weighted sum of pixel values, which we choose to be time-invariant but note that this strong constraint is not necessary in general.

@@

% incorporation of k2sc
While the halo procedure produced a fairly clean light curve in most cases, there were nevertheless residual systematic errors related to spacecraft motion. In order to correct these, we employed the \textsc{k2sc} code \citep{Aigrain2015,k2sc}, which simultaneously models a light curve as a 3D Gaussian Process (GP) in time and predicted position (the K2 standard data product \textsc{pos\_corr}) in pixels $(x,y)$. The model prediction in time for fixed position is then a nonparametric model of the stellar variability, and the prediction for the $x,y$ component evaluated for fixed time represents the pointing systematics. We subtracted the systmatics model from the input fluxes to obtain a final corrected flux, which is the time series we use and recommend for science. Campaigns~9, 10, and~11 were observed in two blocks each, denoted C91/C92, C101/C102 and C111/C112 by the \ktwo Team. The target pixel files for C91, C92, and C101 include no position information, and there are no halo apertures for C112. As a result \textsc{k2sc}-corrected data are not available for these targets, and no halo light curves at all are available in C112.

@@

We calculated halo light curves of 36\,Oph and their CDPPs for $k \in \{1,2,3\}$, and $\delta \in [1,50]$ for long cadence and for various valies of $\delta \in [1,2500]$ for short cadence data. The results are displayed in Figures~\ref{cdpps_lc} and~\ref{cdpps_sc}. We found that for long cadence data, $k=1$ (TV) and a lag $\delta=10$ provide the best CDPP. This is unsurprising: that this is one cadence shorter than the 12~cadence thruster firing period. In this context we can understand the optimum as suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ with slow improvement with larger $\delta$, and performs very poorly for $k=3$ at all lags.

@@

\subsection{Oscillating Red Giants}
\label{sec:rgs}
31 of the evolved stars in our sample have detectable stochastically-excited solar-like acoustic (p-mode) oscillations. In the asymptotic limit, these consist of a comb of modes separated by the large frequency separation \Dnu, approximately the sound-crossing-time of the star, with a Gaussian envelope centred on the frequency of maximum power \numax, which scales with the acoustic cutoff frequency at the star's surface. 
These \Dnu and \numax values can be used to constrain stellar fundamental parameters, such as radius, mass, and age \citep[e.g.][for a recent review]{2017A&ARv..25....1H}. Detailed studies of the deviations from the asymptotic limit for p-modes, e.g. due to acoustic `glitches', provide information on the He content and mixing processes at the bottom of the convective envelope \citep[e.g. ][]{Verma2019}. On the other hand, dipole mixed modes, which have a g-mode character in the inner regions of the star, fulfil an asymptotic period spacing determined by the buoyancy frequency inside the star. This spacing can be used to accurately determine the stellar evolutionary state and allow to distinguish between hydrogen shell and core helium burning \citep{bedding2011}. 
Summary plots for a good example of such a star, $\eta$~Cancri, are shown in Figure~\ref{fig:etacnc}.

Using the Sydney pipeline \citep{Huber2009} with modifications to the extraction of \Dnu detailed in \citet{Yu2018}, we extract the global asteroseismic parameters \numax and \Dnu for the~31 giants for which oscillations are detected with sufficient signal-to-noise. These parameters are listed in Table~\ref{rgs}; the stars are noted as showing `RG' variability in Table~\ref{table_all}, whereas this field is left blank for stars of luminosity class III for which oscillations are not unambiguously detected. High precision spectroscopy of these stars would permit detailed stellar modelling and the extraction of precise elemental abundances, which would make these stars useful as benchmarks for large spectroscopic surveys or testing detailed stellar models. This sample will be an addition to the~36 Gaia FGK benchmark stars \citep{gaiabenchmark1,gaiabenchmark3,2018RNAAS...2c.152J}, the 23 BRITE-Constellation asteroseismic giants \citep{Kallinger2019}, and the~33 \kepler Smear Campaign spectroscopic benchmark red giants \citep{smearcampaign}.

@@

Our dataset includes a rich variety of classical pulsators. We visually inspected the light curves and amplitude spectra to classify all non red-giant stars into traditional variability classes. We identify 23 stars showing $\delta$ Scuti pulsations and 20 with $\gamma$ Doradus pulsations, including 9~with hybrid $\delta$\,Sct/$\gamma$\,Dor variability; 14~slowly pulsating B~stars (SPB stars), 3~$\beta$ Cephei pulsators, and 3~Cepheids; as well as 3~O\,stars and 5~supergiants showing variability \citep[as in][]{Aerts2018,bowman19}. In addition to this, eight stars show rotational modulation, of which two have the characteristics of $\alpha^2$\,CVn chemical spot modulation. The classes we have determined for each star are listed in Table~\ref{table_all}. A detailed frequency analysis of the variability in each star will be presented in a forthcoming paper.

@@

This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute. BJSP also acknowledges the financial support of the Clarendon Fund and Balliol College. TRW acknowledges the support of the Australian Research Council (grant DP150100250) and the Villum Foundation (research grant 10118). SA acknowledges support from the UK Science and Technology Facilities Council (STFC) under grants ST/N000919/1, ST/S000488/1, and ST/R004846/1.

@@

>>>>

\author[0000-0002-0007-6211]{Jie Yu}
\affiliation{Max Planck Institute for Solar System Research, Justus-von-Liebig-Weg 3, 37077 G\"{o}ttingen, Germany}

@@

\author[0000-0003-1822-7126]{Conny Aerts}
\affiliation{Instituut voor Sterrenkunde, KU Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium}
\affiliation{Department of Astrophysics, IMAPP, Radboud University Nijmegen, P.O. Box 9010, NL-6500 GL Nijmegen, The Netherlands}
\affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"{o}nigstuhl 17, D-69117 Heidelberg}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
While the \kepler\ Mission was designed to look at tens of thousands of faint stars ($V \gtrsim 12$), brighter stars that saturated the detector are important because they can be and have been observed very accurately by other instruments. By analyzing the unsaturated scattered-light `halo' around these stars, we have retrieved precise light curves of most of the brightest stars in \ktwo fields from Campaign~4 onwards. The halo method does not depend on the detailed cause and form of systematics, and we show that it is effective at extracting light curves from both normal and saturated stars. The key methodology is to optimize the weights of a linear combination of pixel time series with respect to an objective function. We test a range of such objective functions, finding that \textit{lagged Total Variation}, a generalization of Total Variation, performs well on both saturated and unsaturated \ktwo targets. Applying this to the bright stars across the \ktwo Campaigns reveals stellar variability ubiquitously, including effects of stellar pulsation, rotation, and binarity. We describe our pipeline and present a catalogue of the 161 bright stars, with classifications of their variability, asteroseismic parameters for red giants with well-measured solar-like oscillations, and remarks on interesting objects. These light curves are publicly available as a High Level Science Product from the Mikulski Archive for Space Telescopes (MAST). \href{https://github.com/benjaminpope/k2halo}{\color{linkcolor}\faGithub} % git link shamelessly borrowed from Luger!
\end{abstract}

@@

% intro kepler
The \kepler Space Telescope was launched with a main goal of determining the frequency of Earth-sized planets around Solar-like stars \citep{2010sci...327..977b}. In order to explore these populations it was necessary to observe hundreds of thousands of stars, with the consequence that the \kepler exposure time and gain were set to optimally observe eleventh or twelfth-magnitude stars, while bright stars are saturated and intentionally avoided. In the two-wheeled revival as the \ktwo mission, the \kepler telescope observed a sequence of ecliptic-plane fields containing many more very-saturated stars \citep{2014PASP..126..398H}. While it is difficult to obtain precise light curves of these stars because of their saturation, they are some of the most valuable targets to follow up with photon-hungry methods such as interferometry and high-resolution spectroscopy, and they typically have long histories of previous observations. Dedicated bright-star space photometry missions such as MOST \citep{most} and the BRITE-Constellation \citep{brite,brite2} use very small telescopes (15 and~20\,cm apertures respectively), to assemble time-series photometry of bright stars, but larger telescopes such as \kepler (0.95\,m) lead to higher-precision light curves.

@@

% describe halo algorithm
The `TV-min' halo method was first described by \citet{White2017} and applied to the Pleiades' Seven Sisters. It was also applied to Aldebaran with further developments by \citet{Farr2018}. We follow the Optimized-Weight Linear (OWL) photometry concept described by Hogg \&  Foreman-Mackey (2014, unpublished: preprint  \href{https://github.com/davidwhogg/OWL/}{\nolinkurl{github.com/davidwhogg/OWL/}}) in our assumptions. We assume that a star has a wide PSF sampled by many pixels with different sensitivities, which moves around on the detector within a small region and which may vary to a small extent in time. We assume that our time series consists of many epochs sampled with a nearly even cadence. We do not have metadata describing the spacecraft motion, pixel gains, PSF variations or other noise processes. Because photometry is a linear operation, any estimator of the flux is necessarily a weighted sum of pixel values, which we choose to be time-invariant but note that this strong constraint is not necessary in general.

@@

% incorporation of k2sc
While the halo procedure produced a fairly clean light curve in most cases, there were nevertheless residual systematic errors related to spacecraft motion. In order to correct these, we employed the \textsc{k2sc} code \citep{Aigrain2015,k2sc}, which simultaneously models a light curve as a 3D Gaussian Process (GP) in time and predicted position (the K2 standard data product \textsc{pos\_corr}) in pixels $(x,y)$. The model prediction in time for fixed position is then a nonparametric model of the stellar variability, and the prediction for the $x,y$ component evaluated for fixed time represents the pointing systematics. We subtracted the systematics model from the input fluxes to obtain a final corrected flux, which is the time series we use and recommend for science. Campaigns~9, 10, and~11 were observed in two blocks each, denoted C91/C92, C101/C102 and C111/C112 by the \ktwo Team. The target pixel files for C91, C92, and C101 include no position information, and there are no halo apertures for C112. As a result \textsc{k2sc}-corrected data are not available for these targets, and no halo light curves at all are available in C112.

@@

We calculated halo light curves of 36\,Oph and their CDPPs for $k \in \{1,2,3\}$, and $\delta \in [1,50]$ for long cadence and for various valies of $\delta \in [1,2500]$ for short cadence data. The results are displayed in Figures~\ref{cdpps_lc} and~\ref{cdpps_sc}. We found that for long cadence data, $k=1$ (TV) and a lag $\delta=10$ provide the best CDPP. As this is one cadence shorter than the 12~cadence thruster firing period, we can understand the optimum as suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for $k \in \{1,2\}$ with slow improvement with larger $\delta$, and performs very poorly for $k=3$ at all lags.

@@

\subsection{Oscillating Red Giants}
\label{sec:rgs}
Thirty-one of the red giants in our sample have detectable stochastically-excited solar-like acoustic (p-mode) oscillations. In the asymptotic limit, these consist of a comb of modes separated by the large frequency separation \Dnu, approximately the sound-crossing-time of the star, with a Gaussian envelope centred on the frequency of maximum power \numax, which scales with the acoustic cutoff frequency at the star's surface. 
These \Dnu and \numax values can be used to constrain stellar fundamental parameters, such as radius, mass, and age \citep[e.g.][for a recent review]{2017A&ARv..25....1H}. Detailed studies of the deviations from the asymptotic limit for p-modes, e.g. due to acoustic `glitches', provide information on the He content and mixing processes at the bottom of the convective envelope \citep[e.g. ][]{Verma2019}. On the other hand, dipole mixed modes, which have a g-mode character in the inner regions of the star, fulfil an asymptotic period spacing determined by the buoyancy frequency inside the star. This spacing can be used to accurately determine the stellar evolutionary stage, and allows us to distinguish between hydrogen shell and core helium burning \citep{bedding2011}. 
Summary plots for a good example of such a star, $\eta$~Cancri, are shown in Figure~\ref{fig:etacnc}.

Using the Sydney pipeline \citep{Huber2009} with modifications to the extraction of \Dnu detailed in \citet{Yu2018}, we extract the global asteroseismic parameters \numax and \Dnu for the~31 red giants for which oscillations are detected with sufficient signal-to-noise. These parameters are listed in Table~\ref{rgs}; the stars are noted as showing `RG' variability in Table~\ref{table_all}, whereas this field is left blank for stars of luminosity class III for which oscillations are not unambiguously detected. High precision spectroscopy of these stars would permit detailed stellar modelling and the extraction of precise elemental abundances, which would make these stars useful as benchmarks for large spectroscopic surveys or testing detailed stellar models. This sample will be an addition to the~36 Gaia FGK benchmark stars \citep{gaiabenchmark1,gaiabenchmark3,2018RNAAS...2c.152J}, the 23 BRITE-Constellation asteroseismic red giants \citep{Kallinger2019}, and the~33 \kepler Smear Campaign spectroscopic benchmark red giants \citep{smearcampaign}.

@@

Our dataset includes a rich variety of classical pulsators. We visually inspected the light curves and amplitude spectra to classify all non red-giant stars into traditional variability classes. We identify 23 stars showing $\delta$ Scuti pulsations and 20 with $\gamma$ Doradus pulsations, including 9~with hybrid $\delta$\,Sct/$\gamma$\,Dor variability; 14~slowly pulsating B~stars (SPB stars), 3~$\beta$ Cephei pulsators, and 3~Cepheids; as well as 3~O\,stars and 5~blue supergiants showing low-frequency variability \citep[as in][]{Aerts2018,bowman19}. In addition to this, the light curves of eight stars reveal rotational modulation, of which two have the characteristics of $\alpha^2$\,CVn chemical spot modulation. The classes we have determined for each star are listed in Table~\ref{table_all}. A detailed frequency analysis of the variability in each star will be presented in a forthcoming paper.

@@

@@

This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute. BJSP also acknowledges the financial support of the Clarendon Fund and Balliol College. TRW acknowledges the support of the Australian Research Council (grant DP150100250) and the Villum Foundation (research grant 10118). SA acknowledges support from the UK Science and Technology Facilities Council (STFC) under grants ST/N000919/1, ST/S000488/1, and ST/R004846/1. CA received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement N$^\circ$670519: MAMSIE) and from the KU\,Leuven Research Council (grant C16/18/005: PARADISE).
